---
title: "convSig-vignette"
author: "Antoine Yi-Cheng Tian"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{convSig-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<center><h2> Abstract </h2></center>

This package implements an entirely novel approach for the detection and extraction
of mutational processes that is based on Yun Feng's work 
(convSig project, Oxford University PhD). <br />
Every cancer-affected patient has specific mutational patterns in their genome,
hence detecting, extracting, and characterising such patterns can provide a useful
source of information for the personalisation of cancer treatment.  The methods 
implemented in this package make use of the convolutional filtering techniques found
in computer vision. Our model is centred around the concept of each mutational process 
being represented as a filter, which slides over the entire genome of interest and 
probabilistically induces mutations in sequences that it has a strong rapport with. Our
methods extract and characterise these processes by solving for variables in this 
model using expectation maximisation (either using the ReLU transform or the exponential
transform).


```{r setup}
library(convSig)
```
